---
title: "Magic Functions to Obtain Results from for Loops in R"
author: Koji MAKIYAMA (@hoxo_m)
output: 
  html_document:
    keep_md: true
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

[![Travis-CI Build Status](https://travis-ci.org/hoxo-m/magicfor.svg?branch=master)](https://travis-ci.org/hoxo-m/magicfor)
[![CRAN Version](http://www.r-pkg.org/badges/version/magicfor)](https://cran.r-project.org/package=magicfor)
[![Coverage Status](https://coveralls.io/repos/github/hoxo-m/magicfor/badge.svg?branch=master)](https://coveralls.io/github/hoxo-m/magicfor?branch=master)

## 1. Overview

`for()` is one of the most popular functions in R.
As you know, it is used to create loops.

For example, let's calculate squared values for 1 to 3.

```{r}
for (i in 1:3) {
  squared <- i ^ 2
  print(squared)
}
```

It is very easy.

However, it becomes too much hassle to change such codes to store the result.
You must prepare some containers with correct length for storing the result and change `print()` to assignment statements.

```{r}
result <- vector("numeric", 3) # prepare a container
for (i in 1:3) {
  squared <- i ^ 2
  result[i] <- squared         # change to assignment
}
result
```

Moreover, you may want to store the result as a data frame with the iteration numbers.

```{r}
result <- data.frame(matrix(nrow = 3, ncol = 2))
colnames(result) <- c("i", "squared")
for (i in 1:3) {
  squared <- i ^ 2
  result[i, 1] <- i
  result[i, 2] <- squared
}
result
```

What a bother!

In such or more troublesome situations like that you have to store many variables, the code will grow more complex soon.

The **magicfor** package makes to resolve the problem being kept readability.

Let's just add two lines before the for loop.
First, load the library. Second, call `magic_for()`.
Notice that the main for loop is kept intact.

```{r}
library(magicfor)               # load library
magic_for(print, silent = TRUE) # call magic_for()

for (i in 1:3) {
  squared <- i ^ 2
  print(squared)
}

magic_result_as_dataframe()     # get the result
```

`magic_for()` takes a function name, and then reconstructs `for()` to remember values passed to the specified function in for loops.
We call the reconstruction "magicalization".
Once you call `magic_for()`, as you just run `for()` as usual, the result will be stored in memory automatically.

Here, we are using `magic_result_as_dataframe()` in order to get the stored values.
It is one of the functions to obtain results from "magicalized for loops", and means to take out the results as data.frame.

Even if the number of observed variables increases, you can do it the same way.

```{r}
magic_for(silent = TRUE)

for (i in 1:3) {
  squared <- i ^ 2
  cubed <- i ^ 3
  put(squared, cubed)
}

magic_result_as_dataframe()
```

`put()` is a useful function defined in the package.
It allows to take any number of variables and can display them.

## 2. Installation

You can install the package from GitHub.

```{r eval=FALSE}
install.packages("devtools") # if you have not installed "devtools" package
devtools::install_github("hoxo-m/magicfor")
```

You can use **githubinstall** package to easy install from GitHub.

```{r eval=FALSE}
install.packages("githubinstall") # if you have not installed "githubinstall" package
githubinstall::githubinstall("magicfor")
```

The source code for githubinstall package is available on GitHub at

- https://github.com/hoxo-m/magicfor.

## 3. Details

The package provides the functions as folows:

- `magic_for()`: Magicalize for.
- `magic_free()`: Free magicalization.
- Get results:
    - `magic_result()`: as a list.
    - `magic_result_as_vetor()`: as a vector.
    - `magic_result_as_dataframe()`: as a data.frame.
- `put()`: Display values.

### 3.1 Basics

The main function `magic_for()` magicalize for loops.
"Magicalize" means to change the behavior of `for()` to store values outputted via the target function.

```{r}
library(magicfor)
magic_for()

for (i in 1:3) {
  squared <- i ^ 2
  put(squared)
}
```

The default target function is `put()`. It displays input values, for example:

```{r}
x <- 1
put(x)
```

You can take out the stored values using `magic_result_**()` after the loop finished.

```{r}
magic_result_as_vector()
```

### 3.2 `magic_for()`

`magic_for()` has several options.

Specify the first argument `func`, you can change target function.

```{r}
magic_for(cat)

for (i in 1:3) {
  squared <- i ^ 2
  cat(squared, " ")
}
```

If `progress = TRUE`, show progress bar.

```{r eval=FALSE}
magic_for(progress = TRUE)

for (i in 1:3) {
  squared <- i ^ 2
  put(squared)
}
```

```
#> |=================================================================| 100%
```

If you set `test` a number, the iteration is limited to that number of times.

```{r}
magic_for(test = 2)

for (i in 1:100) {
  squared <- i ^ 2
  put(squared)
}
```

If `silent = TRUE`, target function will be not executed only the values will be stored.

If `temp = TRUE`, the effect of magicalization will be lost after once execution of for loop.

```{r}
magic_for(temp = TRUE)
is_magicalized()

for (i in 1:3) {
  squared <- i ^ 2
  put(squared)
}

is_magicalized()
```

### 3.3 `magic_free()`

You can use `magic_free()` to free magicalizetion.

```{r}
magic_for()
is_magicalized()

magic_free()
is_magicalized()
```

The function also clean the stored values.

```{r}
magic_for(silent = TRUE)

for (i in 1:3) {
  squared <- i ^ 2
  put(squared)
}

magic_result_as_vector()

magic_free()
magic_result_as_vector()
```

### 3.4 `magic_result_**()`

You can use `magic_result_**()` to obtain results from magicalized for loops.

```{r}
magic_for(silent = TRUE)

for (i in 1:3) {
  squared <- i ^ 2
  put(squared)
}
```

`magic_result()` returns results as a list.

```{r}
magic_result()
```

`magic_result_as_vector()` returns results as a vector.

```{r}
magic_result_as_vector()
```

`magic_result_as_dataframe()` returns results as a data.frame.

```{r}
magic_result_as_dataframe()
```

### 3.5 `put()`

`put()` displays input values with high flexibility.

```{r}
x <- 2
y <- 3
put(x)
put(x, y)
put(x, x ^ 2, x ^ 3)
put(x, squared = x ^ 2, cubed = x ^ 3)
```

This flexibility is very useful for **magicfor**.

```{r}
magic_for()

for (i in 1:3) {
  put(x = i, squared = i ^ 2, cubed = i ^ 3)
}

magic_result_as_dataframe(F)
```

## 4. Related Work

- WANTED!
